---
title: "p8105_hw6_mrc2229"
author: "Matthew Curran"
date: "November 19, 2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
library(rvest)
library(viridis)
library(broom)
library(modelr)
library(mgcv)
```

# Problem 1
## Bring in and clean the birthweight data consisting of roughly 4000 children.

```{r}
birthweight_data = read_csv ("./data/birthweight.csv") %>%
  janitor::clean_names() %>%
  rename(
    b_sex=babysex,
    b_headcirc=bhead,
    b_length=blength,
    b_bwt=bwt,
    m_wt_delivery=delwt, 
    fam_income_month=fincome, 
    f_race=frace,
    gest_age_wks=gaweeks, 
    m_age_menarche=menarche, 
    m_hgt=mheight,
    m_age_delivery=momage,
    m_race=mrace,
    births_prior=parity,
    prev_low_bwt=pnumlbw,
    prev_small_gest=pnumsga,
    m_prepreg_bmi=ppbmi,
    m_prepreg_wt=ppwt,
    ave_smoke=smoken,
    m_preg_wt_gain=wtgain
  ) %>%
  mutate(f_race=as.factor(f_race), m_race=as.factor(m_race), b_sex=as.factor(b_sex), malform=as.factor(malform)) 
```

* Variables were renamed to be more descriptive and follow a more consistent structure. 
* Variables for race, sex, and presence of malformations that could affect weight were converted from numeric to factor variables.
* There was no missing data to fill in or remove.
* There are `r nrow (birthweight_data)` observations in the birthweight data set and `r ncol (birthweight_data)` variables.

## My Predictive Model
* I hypothesize that a mother’s weight gain during pregnancy (in pounds), gestational age (in weeks), and the mother's age during delivery are significant predictors for predicting a baby's birthweight.
* Univariate anaylsis was run on all three of these predictors first to see their individual significance before adding them to a multivariable model, with a 5% level of significance. Only individually significant variables are added to the final model.

```{r}
wt_gain_model = lm(b_bwt ~ m_preg_wt_gain, data = birthweight_data)
summary(wt_gain_model)
coef(wt_gain_model)
```
* With a p-value of <2e-16, there is sufficient evidence to suggest mother’s weight gain during pregnancy is a significant predictor of a baby's birthweight.

```{r}
gest_age_wks_model = lm(b_bwt ~ gest_age_wks, data = birthweight_data)
summary(gest_age_wks_model)
coef(gest_age_wks_model)
```
* With a p-value of <2e-16, there is sufficient evidence to suggest gestational age is a significant predictor of a baby's birthweight.

```{r}
m_age_delivery_model = lm(b_bwt ~ m_age_delivery, data = birthweight_data)
summary(m_age_delivery_model)
coef(m_age_delivery_model)
```
* With a p-value of <2e-16, there is sufficient evidence to suggest mother’s age at delivery is a significant predictor of a baby's birthweight.

* All of the proposed predictors were individually significant, so a multivariable linear model was made using all three.
```{r}
my_bw_model = lm(b_bwt ~ m_preg_wt_gain + gest_age_wks + m_age_delivery, data = birthweight_data)
summary(my_bw_model)
coef(my_bw_model)
```

* The null hypothesis is that this model has no predictive value of a baby's birth weight. The alternative hypothesis is that this model has some predictive value of a baby's birth weight.
* This multivariable model was found to have an F-statistic of 401.9 on 3 and 4338 Degrees of freedom with a p-value of < 2.2e-16.
* Based on this F-statistic and p-value, we reject the null hyphothesis and conclude that the mother’s weight gain during pregnancy (in pounds), gestational age (in weeks), and the mother's age during delivery are significant predictors of a baby's birthweight.
* With an R-squared value of 0.2175, 21.75% of the variation in a baby's birthweight can be explained by this linear model, the rest remains unexplained.

## Plot of model residuals against fitted values
```{r}
resid_fit_plot = birthweight_data %>% 
  add_predictions(my_bw_model) %>%
  add_residuals(my_bw_model) %>%
  ggplot(aes(x = resid, y = pred)) + geom_point() + 
  geom_smooth(aes(y = pred))

resid_fit_plot
```

* The above plot shows my birthwieght predicting linear model's residuals against the fitted values from the model. A smooth predictive line was added to the model.

## Comparing my model to two others
* The first model predicts baby birthweight via the direct effects of length at birth and gestational age as predictors.
* The second model uses head circumference, length, sex, and all their interactions (including the three-way interaction) as the predictors.

```{r}
main_eff_model = lm(b_bwt ~ b_length + gest_age_wks, data = birthweight_data)
interact_model = lm(b_bwt ~ b_length + b_sex + b_headcirc + b_length*b_sex + b_length*b_headcirc + b_sex*b_headcirc + b_length*b_sex*b_headcirc, data = birthweight_data)

summary(main_eff_model)
summary(interact_model)
```

* In terms of the cross-validated prediction error, the model with the lowest root mean squared errors (RMSEs) is suggested to work the best.
* My original model has an RMSE of `r rmse(my_bw_model, birthweight_data)`
* The model with the direct effects of length at birth and gestational age has an RMSE of `r rmse(main_eff_model, birthweight_data)`
* The final model with head circumference, length, sex, and all their interactions has an RMSE of `r rmse(interact_model, birthweight_data)`
* The RMSEs are suggestive that both of the comparison models work better than my proposed model, and that the model with head circumference, length, sex, and all their interactions is the best model of all three. However, to get a sense of model stability we really need to iterate this whole process.

```{r}
bw_cv_df = 
  crossv_mc(birthweight_data, 100)

bw_cv_df = 
  bw_cv_df %>% 
  mutate(my_bw_model  = map(train, ~lm(b_bwt ~ m_preg_wt_gain + gest_age_wks + m_age_delivery, data = .x)),
         main_eff_model  = map(train, ~lm(b_bwt ~ b_length + gest_age_wks, data = .x)),
         interact_model  = map(train, ~lm(b_bwt ~ b_length + b_sex + b_headcirc + b_length*b_sex + b_length*b_headcirc + b_sex*b_headcirc + b_length*b_sex*b_headcirc, data = .x))) %>% 
  mutate(rmse_my_bw_model = map2_dbl(my_bw_model, test, ~rmse(model = .x, data = .y)),
         rmse_main_eff_model = map2_dbl(main_eff_model, test, ~rmse(model = .x, data = .y)),
         rmse_interact_model = map2_dbl(interact_model, test, ~rmse(model = .x, data = .y)))

bw_cv_df %>% 
  select(starts_with("rmse")) %>% 
pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

* Now we get a sense of variance in prediction error and can compare prediction error distributions across method and come to the same conclusion as before; the model with head circumference, length, sex, and all their interactions is the best model of all three and has the most predictive accuracy.